---
title       : Enron Data Series 
subtitle    : Get your data into shape
author      : Christina Brady
job         : Data Scientist, World Bank
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---&vcenter
## Workshop Objectives:
  
1. Download Data
2. Explore: Use a R to understand what data is here, what it looks like and how it is organized.
3. Plan: Decide in which format we need the data in order to conduct analysis.
4. Format: Get the data into that format.  

--- &vcenter
## Download the data  
  
  
Download the data [here] (https://www.cs.cmu.edu/~./enron/).

--- &vcenter
## Explore  
  
  
* How many folders are there?
* What do they contain?
* How many subfolders are there?
* What format are the files in?
* What kind of information do they contain? Is it consistent?

--- &vcenter
## Plan  
  
* Who writes to and receives email from whom? (Network analysis) 
* Can we determine what topics are in the emails? (Text Analysis) 
* Can we summarize the emails? (Text Analysis) 
* Can we detect who is involved in fraud from their emails? (Classification)
* Can we detect if a writer is male or female? (Classification)

--- &vcenter
## Format: Network Analysis (1/2)  
For igraph and circlize, you need a dataframe of edges and a data frame of links that look like this:    
###### Edges:
```{r, echo = FALSE, fig.height= 2, fig.width=4}
edges <- data.frame(name = c("Jane Doe", "John Doe"), email = c("jane.doe@enron.com", "john.doe@enron.com"))
links <- data.frame(to = c("jane.doe@enron.com", "princess.leia@enron.com"), 
                           from = c("john.doe@enron.com", "john.doe@enron.com"), 
                           weights = c(4, 10))
#suppressPackageStartupMessages(library(ReporteRs))
#edgesft <- FlexTable(edges, body.par.props = parProperties(padding = 2), header.par.props = parProperties(padding = 2))
#setFlexTableWidths(edgesft, rep(0.75, 2))
edges
```  
###### Links:
```{r, echo = FALSE}
#linksft <- FlexTable(links, body.par.props = parProperties(padding = 2), header.par.props = parProperties(padding = 2))
#setFlexTableWidths(linksft, rep(0.75, 3))
links
```  

--- &vcenter
## Format: Network Analysis (2/2)  
For network analysis, we may also use an adjacency matrix, which looks like this:  
```{r, echo = FALSE}
adjacency <- data.frame(jane_doe = c(0, 1, 1), 
                        princess_leia = c(1, 0, 0), 
                        john_doe = c(2, 2, 1))
rownames(adjacency) <- c("jane_doe", "princess_leia", "john_doe")
adjacency
```  

--- &vcenter
## Format: Text Analysis
For topic modeling or to use the text in classification, we need our documents in a document term matrix (tm package).  

```{r, echo = FALSE}
dtm_ex <- data.frame(assign = c(1, 1, 1), 
                     deal = c(1, 0, 0),
                     financial = c(0, 1, 1),
                     structure = c(0, 0, 0))
row.names(dtm_ex) <- c("email_1", "email_2", "email_3")
dtm_ex
```  

--- &vcenter
## Format: Classification  
For regression, trees, random forests, etc., we need a matrix of features.

```{r, echo = FALSE}
mexample <- data.frame(name = c("Jane Doe", "Princess Leia", "Han Solo"), 
                       fraudster = c(1, 0, 1), 
                       email_topic = c("football", "resistence", "Leia"), 
                       ave_email_length = c(25, 10, 30))
mexample
```  


--- &vcenter
## Resources:
- Quick igraph [example] (https://www.r-bloggers.com/an-example-of-social-network-analysis-with-r-using-package-igraph/). 
- igraph [vignette] (http://www.kateto.net/wp-content/uploads/2016/01/NetSciX_2016_Workshop.pdf).
- circlize vignette
- tm package
- topic models package
